Using numerical modelling to investigate temperature and wave exposure as drivers

### 2020-01-06 morphology_paper.Rmd
updated map and word output

---
title: Using numerical modelling to investigate temperature and wave exposure as drivers
  of macroalgal morphological characteristics
author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
output:
  word_document:
    reference_docx: frontiers_template.docx
  pdf_document:
    fig_caption: yes
  html_document:
    df_print: paged
bibliography: references.bib
papersize: A4
tables: yes
abstract:  '**Abstract**: Macroalgal morphological variation is determined to a large extent by
  a combination of environmental factors, with wave exposure and temperature
  perhaps the main influences as they are key environmental properties to which a species becomes locally adapted.                Macroalgae have shown to exhibit different responses to different magnitudes of exposure to waves, such as reduction
  in overall size and strength increasing traits. In terms of temperature, warmer
  environments have been shown to reduce the overall size of resident and transplanted
  species. However, none of the past studies have identified specific wave and temperature
  metrics responsible for the morphological adaptation macroalgae exhibit. Past research
  has often used simple or two-dimensional models of wave exposure, which do not
  take into account important aspects of the nearshore environment such as wave breaking,
  refraction and diffraction. Furthermore, past studies have often used satellite-derived
  datasets as sources for temperature data; however, such data have been shown to have large bias when 
  applied to the nearshore environment. This study used *in situ* temperature data and wave power
  metrics calculated from a 3D-numerical model to identify specific temperature and wave metrics
  responsible for morphological adaptation of the kelp, *Ecklonia maxima* and *Laminaria pallida*. The results identify wave      exposure as the main driver of morphological adaptation while identifying specific wave metrics.
  Furthermore, the results show differences in wave metrics between species; and between
  deep and shallow populations.'

@@

```{r map study sites, echo=FALSE, fig.height=5, fig.width=5, message=FALSE, warning=FALSE, fig.cap="\\label{fig:map study sites}Map of the study area and sampling sites. The red dot indicates the position of Cape Town for sake of reference."}

library(png)
library(grid)
library(gridExtra)

map <- readPNG("figures/sa_map.png")
grid.raster(map)
```

@@

```{r schematic, echo=FALSE, fig.cap='\\label{fig:schematic}Schematic of *Laminaria pallida* and *Ecklonia maxima* morphology [@Dyer2018-rv].', fig.height=6, fig.width=6}

schematic <- readPNG("figures/kelp_diagram.png")
grid.raster(schematic)
```

@@

library(png)
library(grid)
library(gridExtra)

wp_1 <- rasterGrob(as.raster(readPNG("figures/CapePen_Total.PNG")))
wp_2 <- rasterGrob(as.raster(readPNG("figures/CapePen_Summer.PNG")))
wp_3 <- rasterGrob(as.raster(readPNG("figures/CapePen_Winter.PNG")))
wp_4 <- rasterGrob(as.raster(readPNG("figures/CapePen_Spring.PNG")))
wp_5 <- rasterGrob(as.raster(readPNG("figures/CapePen_Autumn.PNG")))
wp_6 <- rasterGrob(as.raster(readPNG("figures/CapePen_Legend.PNG")))

>>>>

---
title: Using numerical modelling to investigate temperature and wave exposure as drivers
  of morphological characteristics of the kelps, Ecklonia maxima and Laminaria pallida
author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
output:
  word_document:
    reference_docx: frontiers_template.docx
  pdf_document:
    fig_caption: yes
  html_document:
    df_print: paged
bibliography: references.bib
papersize: A4
tables: yes
abstract:  '**Abstract**: Macroalgal morphological variation is determined to a large extent by
  a combination of environmental factors, with wave exposure and temperature
  perhaps the main influences as they are key environmental properties to which a species becomes locally adapted.                Macroalgae have shown to exhibit different responses to different magnitudes of exposure to waves, such as reduction
  in overall size and strength increasing traits. In terms of temperature, warmer
  environments have been shown to reduce the overall size of resident and transplanted
  species. However, none of the past studies have identified specific wave and temperature
  metrics responsible for the morphological adaptation macroalgae exhibit. Past research
  has often used simple or two-dimensional models of wave exposure, which do not
  take into account important aspects of the nearshore environment such as wave breaking,
  refraction and diffraction. Furthermore, past studies have often used satellite-derived
  datasets as sources for temperature data; however, such data have been shown to have large bias when 
  applied to the nearshore environment. This study used *in situ* temperature data and wave power
  metrics calculated from a 3D-numerical model to identify specific temperature and wave metrics
  responsible for morphological adaptation of the kelp, *Ecklonia maxima* and *Laminaria pallida*. The results identify wave      exposure as the main driver of morphological adaptation while identifying specific wave metrics.
  Furthermore, the results show differences in wave metrics between species; and between
  deep and shallow populations.'

@@

```{r map processing, message=FALSE, warning=FALSE, include=FALSE}
## Load software:
library(ggplot2)
library(rgdal)
library(tidyverse)
library(viridis)
library(gridExtra)
source("map/scaleBarFunc.R")
## Coastline of Southern Africa
load("map/south_africa_coast_hi_res.Rdata")
## Read in metadata
study_sites <- read.csv("map/metadata_map.csv", sep = ";")

study_sites_edit <- study_sites %>%
  filter(site == "St Helena Bay"|site == "Betty's Bay")

## Make first map of whole study area
map_A <- ggplot() +
geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
            size = 1.0, colour = "black", fill = NA) +
  geom_point(data = study_sites, aes(x = lon, y = lat), color = "red") +
  geom_text(data = study_sites_edit, aes(x = lon, y = lat, label = site), nudge_x = -0.1, nudge_y = -0.01) +
  annotate("text", label = "WARM\nTEMPERATE", x = 19.0, y = -34.5, size = 4.0, angle = 0, colour = "black") +
  annotate("text", label = "COOL\nTEMPERATE", x =17.8, y = -33.5, size = 4.0, angle = 0, colour = "black") +
  annotate("text", label = "Cape\nPeninsula", x = 18.4, y = -34.5, size = 3.0, angle = 0, colour = "black") +
  scaleBar(lon = 17.8, lat = -34.45, distanceLon = 10, distanceLat = 10, distanceLegend = -10, dist.unit = "km",
  arrow.length = 10, arrow.distance = 15, arrow.North.size = 3) +
  coord_cartesian(xlim = c(17.75, 19.15), ylim = c(-32.50, -34.55)) +
  theme(axis.ticks = element_blank()) + # Remove tick marks
  xlab("Latitude") +
  ylab("Longitude")
ggsave(filename = "map_A.jpg", plot = last_plot(), device = "jpeg", path = "map/")
## Make 2nd map of just the peninsula
map_B <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  geom_point(data = study_sites, aes(x = lon, y = lat, colour = "red"), show.legend = FALSE) +
  geom_text(data = study_sites, aes(x = lon, y = lat, label = site), nudge_x = 0.02, nudge_y = 0) +
  coord_cartesian(xlim = c(18.275, 18.525), ylim = c(-34.375, -34.100)) +
  theme(axis.ticks = element_blank(), # Remove tick marks
        axis.text = element_blank(), # Remove lat/ lon numbers
        axis.title = element_blank()) # Remove lat/ lon labels
ggsave(filename = "map_B.jpg", plot = last_plot(), device = "jpeg", path = "map/")

# Convert the figures to grobs
grob_A <- ggplotGrob(map_A)
grob_B <- ggplotGrob(map_B)

## Create combined map
map_combined <- ggplot() +
  # First set the x and y axis limits so we know what the ranges are
  coord_equal(xlim = c(1, 10), ylim = c(1, 10), ratio = 1, expand = TRUE, clip = "on") +
  # Then we place our facetsover one another using the coordinates we created
  annotation_custom(grob_A,
                    xmin = 1, xmax = 10, ymin = 1, ymax = 10) +
  annotation_custom(grob_B,
                    xmin = 4, xmax = 10, ymin = 4.5, ymax = 10)
ggsave(filename = "map_combined.jpg", plot = last_plot(), device = "jpeg", path = "map/", width = 8, height = 8)
```

```{r map study sites, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, fig.cap="\\label{fig:map study sites}Map of the study area and sampling sites. The red dot indicates the position of Cape Town for sake of reference."}
map_combined
```

@@

```{r schematic, echo=FALSE, fig.cap='\\label{fig:schematic}Schematic of *Laminaria pallida* and *Ecklonia maxima* morphology [@Dyer2018-rv].', fig.height=6, fig.width=6}
library(jpeg)
library(grid)
library(gridExtra)

schematic <- readJPEG("figures/kelp_diagram.jpg")
grid.raster(schematic)
```

@@

library(jpeg)
library(grid)
library(gridExtra)
library(ggpubr)

wp_1 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6A.jpg")))
wp_2 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6B.jpg")))
wp_3 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6C.jpg")))
wp_4 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6D.jpg")))
wp_5 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6E.jpg")))
wp_6 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6_legend.jpg")))
